Generating Hubbard Model Solutions from Anderson Impurity Model Solutions

### 2021-09-05 notes.tex
reverted to simple fit with 3 params

\title{Generating Hubbard Model Solutions from Anderson Impurity Model Solutions}
\author{Abhirup Mukherjee, Dr. Siddhartha Lal}
\begin{document}
\maketitle
\section{Introduction}
This is an attempt to obtain various quantities like Greens functions, self-energies, spectral functions and (if possible) energies and wavefunctions of the Hubbard model, using a cluster-bath approach. The cluster-bath system is taken to be a single-impurity Anderson model with a correlated bath. The correlation will be brought about in two ways: a self-energy $\Sigma(k,\omega)$ of the bath, and a double occupancy repulsion cost $U_b$. The Hubbard and the correlated single-impurity Anderson models are defined using the Hamiltonians
\begin{align}
H_\text{hubb} &= -t^H\sum_{\sigma,\left<i,j \right>}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + U^H\sum_i \hat n_{i \uparrow} \hat n_{i \downarrow} - \mu^H \sum_{i\sigma}\hat n_{i\sigma}\\
H_\text{siam} &= \sum_{k\sigma}\left[\epsilon_k + \Sigma(k,\omega)\right]\hat n_{k\sigma} + \epsilon_d^A \sum_\sigma\hat n_{d\sigma} + U^A \hat n_{d \uparrow} \hat n_{d \downarrow} + U_b \sum_{kk^\prime}\hat n_k \hat n_{k^\prime} -t^A\sum_{k\sigma}\left(c^\dagger_{d\sigma}c_{k\sigma} + \text{h.c.}\right) \label{clus_bath_siam}
\end{align}
Broadly speaking, the method involves first solving the SIAM using a unitary renormalisation group approach, to get the low energy effective theory, and then combining the low energy Hamiltonians in a symmetrized fashion to get the Hamiltonian for the Hubbard model lattice. It is reminescent of dynamical mean-field theory (DMFT) - both involve an impurity-solver that solves an auxiliary system. The difference, however, lies in the following points:
\begin{itemize}
	\item While DMFT primarily works with Greens functions and self-energies, this method involves Hamiltonians. The impurity-solver in DMFT provides an impurity Greens function (which is then equated with the local Greens function of the bath), while the impurity-solver in this method actually provides a low energy Hamiltonian.
	\item The final step of DMFT is the self-consistency equation, where the impurity and bath-local quantities are set equal. This ensures all sites, along with the impurity site, have the same self-energy, something which is required on grounds of  translational invariance. The present method, however, brings about the translational invariance in a different way. It symmetrizes the Hamiltonians itself, such that all quantites then derived from the Hamiltonian are then guaranteed to have the symmetry.
\end{itemize}
The meaning of each of these statements will become clearer when we describe the method in more detail.

@@

>>>>

\title{Generating Hubbard Model Solutions from Anderson Impurity Model Solutions}
\author{Abhirup Mukherjee, Dr. Siddhartha Lal}
\begin{document}
\maketitle
\tableofcontents
\newpage
\section{Introduction}
This is an attempt to obtain various quantities like Greens functions, self-energies, spectral functions and (if possible) energies and wavefunctions of the Hubbard model, using a cluster-bath approach. The cluster-bath system is taken to be a single-impurity Anderson model with a correlated bath. The correlation will be brought about in two ways: a self-energy $\Sigma(k,\omega)$ of the bath, and a double occupancy repulsion cost $U_b$. The Hubbard and the correlated single-impurity Anderson models are defined using the Hamiltonians
\begin{align}
H_\text{hubb} &= -t^H\sum_{\sigma,\left<i,j \right>}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + U^H\sum_i \hat n_{i \uparrow} \hat n_{i \downarrow} - \mu^H \sum_{i\sigma}\hat n_{i\sigma}\\
H_\text{siam} &= \sum_{k\sigma}\left[\epsilon_k + \Sigma(k,\omega)\right]\hat n_{k\sigma} + \epsilon_d^A \sum_\sigma\hat n_{d\sigma} + U^A \hat n_{d \uparrow} \hat n_{d \downarrow} + U_b \sum_{kk^\prime}\hat n_k \hat n_{k^\prime} -t^A\sum_{k\sigma}\left(c^\dagger_{d\sigma}c_{k\sigma} + \text{h.c.}\right) \label{clus_bath_siam}
\end{align}
Broadly speaking, the method involves first solving the SIAM using a unitary renormalisation group approach, to get the low energy effective theory, and then combining the low energy Hamiltonians in a symmetrized fashion to get the Hamiltonian for the Hubbard model lattice. It is reminescent of dynamical mean-field theory (DMFT) - both involve an impurity-solver that solves an auxiliary system. The difference, however, lies in the following points:
\begin{itemize}
	\item While DMFT primarily works with Greens functions and self-energies, this method involves Hamiltonians. The impurity-solver in DMFT provides an impurity Greens function (which is then equated with the local Greens function of the bath), while the impurity-solver in this method actually provides a low energy Hamiltonian.
	\item The final step of DMFT is the self-consistency equation, where the impurity and bath-local quantities are set equal. This ensures all sites, along with the impurity site, have the same self-energy, something which is required on grounds of  translational invariance. The present method, however, brings about the translational invariance in a different way. It symmetrizes the Hamiltonians itself, such that all quantites then derived from the Hamiltonian are then guaranteed to have the symmetry.
\end{itemize}
The meaning of each of these statements will become clearer when we describe the method in more detail.

@@

\section{Spectral function of the single-impurity Anderson model}
To get a better look at what we are getting into and what we should expect going forward, we first calculate the spectral function of the smaller single-impurity Anderson model that is obtained as the low-energy Hamiltonian of an URG analysis of the full SIAM. This ow energy effective Hamiltonian is
\begin{equation}\begin{aligned}
	\epsilon_d \hat n_d + U\hat n_{d \uparrow} \hat n_{d \downarrow} + v \left(c^\dagger_{d\sigma}c_{0\sigma} + \text{h.c.}\right) + \sum_{k \in \left[-\Lambda^*, \Lambda^*\right], \sigma}\epsilon_k \hat n_{k\sigma}
\end{aligned}\end{equation}
All the impurity couplings are fixed point values, and the momentum states of the bath now run over a reduced bandwidth. The subscript \(d\) indicates that the operator is that of the impurity, while the subscript \(0\) indicates it is of the zeroth site of the bath: \(c_0 = \frac{1}{\sqrt N}\sum_k c_k\). The spectral function of the SIAM is already known from other methods like the numerical renormalization group. For small values of \(U\), the spectral function consists of just the central quasiparticle excitation peak. At larger values of \(U\), we see the emergence of the two Hubbard side bands while the spectral weight at the centre decreases. We also calculated the spectral function of the low energy theory sing exact diagonalization. The plots are shown in fig.~\ref{siam_spec_func}.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.9\textwidth]{./siam_specfunc_all.pdf}
	\caption{Spectral function of the effective SIAM Hamiltonian for various values of \(\frac{U}{\Delta}\). \(\Delta\) is defined by the height of the central peak in the non-interacting model \(U=0\). The inset shows a zoom-in of the final spectral function \(\left( \frac{U}{\Delta} = 19.71 \right)\) near \(\omega = 0\).}
	\label{siam_spec_func}
\end{figure}
\\\\
Because we kept only a discrete bath of 4 momentum states, \textit{we had to artificially enforce the constancy of the height of the central peak.} To convert the delta functions into a continuous spectrum, we  The central peak (or equivalently, the central pole in the impurity Greens function) does not vanish even for exhorbitantly large values of \(\frac{U}{\Delta}\), implying that there is no stable local moment phase in the SIAM (a local moment phase would involve a gap at zero frequency in the impurity spectral function). Formally, the local moment phase will be stable only at \(U \to \infty\) or \(v^* = 0\).
\\\\
We also checked how the width of the central peak varies with increasing \(U\). The behaviour is shown in fig.~\ref{fwhm}. The width was also fitted against a function, as shown in the plot.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.6\textwidth]{./fwhm_vs_U_hrk_fit.pdf}
	\caption{The decay of the full width at half maximum (FWHM) of the central peak, with increasing \(\frac{U}{\Delta}\). The behaviour satisfies a function of the form shown in the legend.}
	\label{fwhm}
\end{figure}

\section{Spectral function of a correlated SIAM}
Since the SIAM cannot show a transition from a metallic (Fermi liquid) phase to a local moment phase, we next consider the slightly more correlated version where the zeroth site (which the impurity couples to) now has a Hubbard repulsion.
\begin{equation}\begin{aligned}
	\epsilon_d \hat n_d + U\hat n_{d \uparrow} \hat n_{d \downarrow} + U_b\hat n_{0 \uparrow} \hat n_{0 \downarrow} + v \left(c^\dagger_{d\sigma}c_{0\sigma} + \text{h.c.}\right) + \sum_{k \in \left[-\Lambda^*, \Lambda^*\right], \sigma}\epsilon_k \hat n_{k\sigma}
\end{aligned}\end{equation}
In such a model, if the impurity is to hybridize with the bath, it has to hop onto the zeroth site, but that site now has a double-occupancy cost. As a result, this new physics makes it harder for the impurity electron to delocalise, and suggests that there might be a metal-insulator transition at some critical value of the impurity \(U\), and this critical value should itself be a function of \(U_b\). This model therefore provides a tool for showing the changes in the phase diagram of the SIAM. Without any correlation in the bath, the SIAM has just one stable fixed point, the strong-coupling one. Starting up a correlation on the zero mode allows us to examine how this scenario might change if we have a correlated bath. One can also draw a comparison to DMFT and argue that such a model is similar to that stage of DMFT when the process has gone through a few steps of self-consistency enforcement and the bath itself has started to become correlated.
\\\\
Apart from its importance as a standalone problem, this correlated SIAM also sets the stage for a fully-symmetrized auxiliary model, the Hubbard dimer in a bath. In other words, by taking a symmetric combination of the correlated SIAM Hamiltonians (the symmetrization is done by exchanging the impurity and bath zero mode sites), one can obtain the Hamiltonian of a dimer hybridising with a bath, as shown in eq.~\ref{dimer_p_bat}. Such a symmetrised Hamiltonian should presumably correspond to a DMFT final-stage Hamiltonian, because the impurity and bath zero mode sites are now completely symmetric.
